Speed and strength of an epidemic intervention

### 2020-10-13 speedstrength.tex
comments on the review & new title
power

\begin{flushleft}{
	\Large
	\textbf\newline{
		Speed and strength of an epidemic intervention
	}
}
\newline
\\
Jonathan Dushoff\textsuperscript{1,2,3,*}
Sang Woo Park\textsuperscript{4}
\\
\bigskip
\textbf{1} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\textbf{2} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{3} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{4} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\bigskip

@@

@@

@@

The importance of speed-based perspectives are still rarely recognized in the case of infectious disease, however. 
Responses to the 2014 Ebola Outbreak in West Africa and the recent COVID-19 outbreak show an over-emphasis on strength at the expense of speed:   
during the early phases of both outbreaks, many disease modelers tried to estimate $\mathcal R_0$ but overlooked $r$.
For example, only 1 out of 7 preliminary analyses of the COVID-19 outbreak that were published as preprints between January 23--26, 2020 reported the doubling time of an epidemic \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
We suggest that infectious disease modelers should be aware of the complementarity of these two frameworks when analyzing disease outbreaks.

>>>>

\begin{flushleft}{
	\Large
	\textbf\newline{
		Strength and speed of an epidemic intervention
	}
}
\newline
\\
Jonathan Dushoff\textsuperscript{1,2,3,*}
Sang Woo Park\textsuperscript{4}
\\
\bigskip
\textbf{1} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\textbf{2} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{3} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{4} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\bigskip

@@

\swp{I also changed the title to strength and speed}

@@

\swp{A new suggested paragraph. I think some version of this paragraph may help deal with above sentence (the image you sent me saying that you don't understand). I think it can also help deal with the reviewer's comments about the growth phase}
\swp{
Here, we focused on the initial growth phase an epidemic, assuming that the observed growth rate \rr is fixed throughout our analysis.
For endemic diseases, as discussed previously, strength-based framework can provide a clearer perspective---therefore, it makes more sense to rely on fixed \RR when evaluating the difficulty of elimination.
In these cases, parameter uncertainty will generally not affect the efficacy of strength-like interventions but can affect conclusions about speed-lik interventions:
given \RR, shorter generation intervals imply faster epidemic \citep{eaton2014proportion, powers2014impact, park2019practical}, requiring faster intervention to control .
While both strength- and speed-based frameworks will give the same answers for a given set of assumptions about an epidemic,
these conclusions can depend on whether we calibrate our models to the strength or speed of an epidemic.
}

@@

The importance of speed-based perspectives are still rarely recognized in the case of infectious disease, however. 
Responses to the 2014 Ebola Outbreak in West Africa and the recent COVID-19 outbreak show an over-emphasis on strength at the expense of speed:   
during the early phases of both outbreaks, many disease modelers tried to estimate $\mathcal R_0$ but overlooked $r$.
For example, only 1 out of 7 preliminary analyses of the COVID-19 outbreak that were published as preprints between January 23--26, 2020 reported the doubling time of an epidemic \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. \swp{does this feel weak now?}
We suggest that infectious disease modelers should be aware of the complementarity of these two frameworks when analyzing disease outbreaks.
